---
title: "Eigenvectors from Eigenvalues"
author: "Jesús M. Castagnetto"
date: "2019-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Motivation

There is a very interesting new result that indicates that the
eigenvectors for a Hermitian (symmetric) matrix, can be calculated from the eigenvalues ["EIGENVECTORS FROM EIGENVALUES" PETER B. DENTON, STEPHEN J. PARKE, TERENCE TAO, AND XINING ZHANG](https://arxiv.org/pdf/1908.03795.pdf)

In this document, I am trying to (kludgily) repeat, using **R**, the excercise that Corey Chivers did in Python at: http://predictivehealthcare.pennmedicine.org/Eigenvectors%20from%20Eigenvalues.html

## Step 1: do the calculations with R's matrix operations

Let's make our matrix `x` and make sure is Hermitian (symmetric)

```{r}
set.seed(20191121)
n = 5
N = n*n
x <- rnorm(N)
dim(x) <- c(n, n)
for(i in 1:n){
  for (j in i:n) {
    x[j, i ] <- x[i, j]
  }
}
isSymmetric(x)
x
```

Now, we will calculate the eigenvectors and eigenvalues the usual way

```{r}
eig <- eigen(x)
eig$values
(result <- abs(eig$vectors))
```

## Step 2: Let's calculate using the calculations from the paper

We prepare the vectors and intermediate matrix needed for the calculations

```{r}
# the eigenvalues of x
lambda_a <- eig$values
# an appropriate matrix
lambda_m <- matrix(nrow = n, ncol = n - 1)
# index list
tidx <- 1:5
for (j in 1:n) {
  idx <- tidx[-j]
  Mj <-  x[idx,idx]
  lambda_m[j,] <-  eigen(Mj)$values
}
```

Then we estimate de eigenvectors

```{r}
v_norm <- matrix(data = 0, nrow = n, ncol = n)
for(i in 1:n) {
  for(j in 1:n) {
    numerator <- 1
    denominator <- 1
    for(k in 1:(n - 1)){
      numerator <- numerator * (lambda_a[i] - lambda_m[j, k])
    }
    for(k in 1:n) {
      if (k != i) {
        denominator <- denominator * (lambda_a[i] - lambda_a[k])
      }
    }
    v_norm[i, j] <- numerator / denominator
  }
}
# the eigenvectors using the new result from the paper
new_result <- t(sqrt(v_norm))
```

## Step 3: Let's compare them

```{r}
# the original eigenvector in absolute value
result
# and the one made using the new formula
new_result
```

They seem to be about the same, let's use the same tolerance as Corey's:

```{r}
abs(result - new_result) < 1e-10
```

Cool!

[@jmcastagnetto](https://twitter.com/jmcastagnetto)